Illustrating Model Properties

### 2024-05-30 model.qmd
improve beta pdfs with unknowns

```{python}
#| label: fig-model-prevalences-with-ipsi
#| fig-cap: "Comparison of predicted (histograms) vs observed (beta posteriors) prevalences. Shown are four scenarios, all including contralateral LNL II involvement: Early T-category and an ipsilateral N0 neck (green), early T-category and ipsilateral LNL II involvement (blue), as well as the same two scenarios but for advanced T-category (orange and red). The figure shows that the model is capable of describing the correlation between ipsi- and contralateral involvement. Although for the scenario of LNL II involvement in both sides, the prediction's split between early and advanced T-category is not large enough."
import matplotlib.pyplot as plt
from lyscripts.scenario import Scenario
from lyscripts.plot.utils import Histogram, BetaPosterior, draw, COLORS
from scripts import paths, shared

fig, ax = plt.subplots()
content = []
colors = iter([COLORS[c] for c in ["orange", "green", "red", "blue"]])

for t_stage in ["early", "late"]:
  for diagnosis in [{"II": True}, {lnl: False for lnl in ["I", "II", "III", "IV", "V"]}]:
    color = next(colors)
    scenario = Scenario(
      t_stages=[t_stage],
      diagnosis={
        "ipsi": {"max_llh": diagnosis},
        "contra": {"max_llh": {"II": True}},
      }
    )
    content.append(
      Histogram.from_hdf5(
        filename=paths.model_dir / "full" / "prevalences_with_ipsi.hdf5",
        dataname=scenario.md5_hash("prevalences"),
        color=color,
        label="prediction"
      )
    )
    beta_post = BetaPosterior.from_hdf5(
      filename=paths.model_dir / "full" / "prevalences_with_ipsi.hdf5",
      dataname=scenario.md5_hash("prevalences"),
      color=color,
    )
    beta_post.kwargs["label"] = f"data: {beta_post.num_success} of {beta_post.num_total}"
    content.append(beta_post)

draw(ax, content, xlims=(0., 25.))
ax.set_title("Observed vs Predicted Prevalence\nof Contralateral LNL II Involvement")
ax.set_yticks([])
ax.set_xlabel("prevalence [%]")

shared.split_legends(
  ax=ax,
  titles=["early with ipsi II", "early ipsi N0", "late with ipsi II", "late with ipsi N0"],
  locs=[(0.23, 0.65), (0.1, 0.99), (0.6, 0.65), (0.45, 0.99)],
)

>>>>

```{python}
#| label: fig-model-prevalences-with-ipsi
#| fig-cap: "Comparison of predicted (histograms) vs observed (beta posteriors) prevalences. Shown are four scenarios, all including contralateral LNL II involvement: Early T-category and an ipsilateral N0 neck (green), early T-category and ipsilateral LNL II involvement (blue), as well as the same two scenarios but for advanced T-category (orange and red). The figure shows that the model is capable of describing the correlation between ipsi- and contralateral involvement. Although for the scenario of LNL II involvement in both sides, the prediction's split between early and advanced T-category is not large enough."
import matplotlib.pyplot as plt
from matplotlib import gridspec
import numpy as np
from lyscripts.scenario import Scenario
from lyscripts.plot.utils import Histogram, BetaPosterior, draw, COLORS
from scripts import paths, shared
from scripts.marginalize_midext import (
  compute_possible_observed_prevalences,
  SummedBetaPosterior,
)

full.set_title(
  "Observed vs Predicted Prevalence of Contralateral LNL II Involvement\nDependent on Ipsilateral Involvement",
  pad=20.,
  fontsize="large",
)

for i, t_stage in enumerate(["early", "late"]):
  axes[i,0].set_ylabel(f"{t_stage} T-cat.", fontweight="bold")
  for j, midext in enumerate([False, True]):
    axes[0,j].set_title("Mid. ext." if midext else "Lateralized", fontweight="bold")
    axes[0,j].set_xticks([])
    axes[-1,j].set_xlabel("prevalence [%]")
    axes[i,j].set_yticks([])
    content = []
    for ipsi, color in zip(
      [{"I": False, "II": False, "III": False, "IV": False, "V": False}, {"III": True}],
      [COLORS["blue"], COLORS["orange"]],
    ):
      scenario = Scenario(
        t_stages=[t_stage],
        midext=midext,
        diagnosis={
          "ipsi": {"max_llh": ipsi},
          "contra": {"max_llh": {"II": True}},
        }
      )
      content.append(
        Histogram.from_hdf5(
          filename=paths.model_dir / "full" / "prevalences_with_ipsi.hdf5",
          dataname=scenario.md5_hash("prevalences"),
          color=color,
          label="prediction",
        )
      )
      beta_post = BetaPosterior.from_hdf5(
        filename=paths.model_dir / "full" / "prevalences_with_ipsi.hdf5",
        dataname=scenario.md5_hash("prevalences"),
        color=color,
      )
      beta_post.kwargs["label"] = f"data: {beta_post.num_success} of {beta_post.num_total}"
      content.append(beta_post)
    draw(
      axes=axes[i,j],
      contents=content,
      xlims=(0., 30. if midext else 10.),
    )

# manually place legends
shared.split_legends(
  ax=axes[0,0],
  titles=["ipsi N0", "ipsi LNL III"],
  locs=[(0.3, 0.99), (0.45, 0.55)],
)
shared.split_legends(
  ax=axes[0,1],
  titles=["ipsi N0", "ipsi LNL III"],
  locs=[(0.3, 0.99), (0.5, 0.55)],
)
shared.split_legends(
  ax=axes[1,0],
  titles=["ipsi N0", "ipsi LNL III"],
  locs=[(0.2, 0.99), (0.55, 0.99)],
)
shared.split_legends(
  ax=axes[1,1],
  titles=["ipsi N0", "ipsi LNL III"],
  locs=[(0.2, 0.99), (0.65, 0.99)],
)
